\typeout{NT FILE chapter-conclusion-retrospective.tex}%
\chapter{Conclusion and Retrospective}
\label{cha:conclusion-retrospective}
This chapter summarizes the work conducted during this internship, highlighting
the key achievements, challenges faced, and lessons learned. It also provides a
retrospective analysis of the project, reflecting on the development process
and the overall impact of the work.

\section{Summary of Achievements}

The work carried out during this internship resulted in two complementary
software systems, designed and implemented in C++ over the course of the
internship period.

The \gls{sdr-control-system} provides a higher-level abstraction over the
\gls{DTAPI}, exposing device management, channel configuration, and signal
transmission functionality through a \gls{gRPC} interface. This addresses
the original problem identified at the outset of the internship: that
configuring \gls{SDR} hardware directly through the \gls{DTAPI} required
detailed knowledge of low-level hardware concerns, such as FIFO buffer sizing
and modulation parameter composition, which the developed system abstracts
away behind a clean, strongly-typed, and remotely accessible interface. The
client-server architecture, together with \gls{gRPC}'s reflection
capabilities, additionally enables interactive use through a \gls{CLI}
client, lowering the barrier for engineers unfamiliar with the underlying
hardware to issue commands and inspect available functionality.

The \gls{rs232-interface} provides a lightweight communication layer over
\gls{RS232}, built around a protocol registration model that allows new
message types and handlers to be declared without modifying the core of the
system. Its design, relying on Boost.PFR for automatic struct serialisation
and \gls{COBS} for packet framing, achieves a level of extensibility and
ease of use comparable to that provided by \gls{gRPC} in the
\gls{sdr-control-system}, while remaining lightweight enough to be
appropriate for a low-level serial link.

Both systems share a consistent set of design principles established during
the internship: the use of errors as values rather than exceptions, the
adoption of XMake as the build system, and a preference for minimal,
composable abstractions over heavier frameworks where the simpler approach
was sufficient. Beyond the two systems themselves, the work also included
their integration with existing proprietary systems at \gls{GMV}, as described in
\autoref{cha:integration-testing}, validated through a combination of unit
testing, code coverage analysis, and the external tester component developed
to simulate measurement equipment in the prototype environment.

\section{Challenges Faced}

Several challenges were encountered throughout the internship, spanning both
technical implementation details and the evolving scope of the project
itself.

On the technical side, the development of the \gls{rs232-interface} surfaced
a number of subtle issues that required careful debugging. The serialisation
layer relies on a trait-based pattern, \texttt{CustomSerde<T>}. This
revealed a heap corruption issue, ultimately traced to deserialization returning a reference instead of the actual deserialized object/struct, which for those unfamiliar with C++'s value semantics can be a common pitfall.
Similarly, the \gls{COBS} implementation required correcting an off-by-one error in output buffer
sizing and an incorrect use of \texttt{pop\_back()} in the decode path,
both of which only manifested under specific payload lengths and were not
immediately obvious from initial testing.

Hardware-level challenges were also significant, particularly around serial
port configuration. The WCH CH382 PCIe card initially relied on the generic
\texttt{8250} driver, which proved insufficient for reliable operation and
was replaced with the vendor-supplied \texttt{ch35\_38x\_linux} driver,
exposing the expected \texttt{/dev/ttyWCH0--3} device nodes. Further issues
were traced to incorrect \texttt{termios} configuration, including the
handling of the \texttt{VMIN} parameter. Another quite relevant issue was the use of 2 \gls{DTE} devices connected back-to-back without a null modem adapter, which, in simple terms, 
meant a device transmitted data into the other's transmit line instead of its receive line. 
This was resolved by using a null modem adapter, which correctly cross-connected the transmit and receive lines between the two devices.

\section{Conclusion}

The internship achieved its core objectives: a \gls{gRPC}-based abstraction
over the \gls{DTAPI} was designed and implemented, providing a programmatic
and remotely accessible interface for \gls{SDR}-based \gls{GNSS} signal
transmission, and a complementary \gls{RS232} communication layer was
developed to support integration with external equipment. Together, these
systems address the gap identified in \autoref{cha:state-of-the-art} between
low-level, hardware-specific vendor APIs and the higher-level, programmatic
control required for automated and remotely operated test workflows.

Beyond the specific technical deliverables, the internship provided practical
exposure to the realities of developing software against real, professional
hardware. The need to adapt the original work plan in
response to a changed protocol requirement and an expanded integration scope
also reflected a broader lesson about software engineering in a professional
context: that requirements and scope are rarely fixed in advance, and that
designing systems with extensibility and clear interfaces in mind, as was
done throughout this work, pays off when that flexibility is inevitably
required.

Overall, the work developed during this internship is believed to represent
a meaningful and reusable contribution within \gls{GMV}, providing a foundation
that can be extended to support additional \gls{SDR} hardware features,
additional \gls{RS232}-based protocols, or further integration with other
systems, beyond what was required within the scope of this internship.