%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% chapter1.tex
%% NOVA thesis document file
%%
%% Chapter with introduction
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\typeout{NT FILE chapter3.tex}%

\chapter{SDR Control}
\label{cha:sdr-control}

\prependtographicspath{{\DIRFIGURES/Covers/}}
\textcolor{Red}{Change this to SDR Facade or Interface}

This chapter describes the \gls{SDR} control system used in this internship.
The \gls{SDR} control system is responsible for managing the communication
between the \gls{SDR} hardware and the software applications that utilize it.
It provides an interface for configuring the \gls{SDR} parameters, such as
frequency, rf-control, and sampling rate, as well as handling data transmission
and reception.

\section{Objective and Motivation}

One of the main problems in this internship, whilst developing the \gls{SDR}
control, was the lack of an easy to use interface. Whilst \gls{DTAPI}, the
\gls{SDR} control library used in this internship, provides a powerful set of
instructions to control the \gls{SDR} hardware, its interface is not
user-friendly and does not offer a clean way to manage the \gls{SDR}
parameters. To address this issue, the \gls{SDR} control system was designed to
provide a more intuitive and user-friendly interface for managing parameters
and operations, either by providing easy to use abstractions to transmit
signals, or by providing synchronization mechanisms between devices.

\section{Design and Implementation}

The \gls{SDR} control system was implemented using a modular approach, with
separate components responsible for device management, signal playback,
synchronization, etc.. A design goal internship was to integrate this system
with other components, \textcolor{Red}{that will not be described here due to
    the scope of this report and their proprietary nature}, to provide a complete
solution for \gls{GNSS} signal generation and transmission. Thus the system was
designed with a client server architecture, where the control system acts as a
server that listens for commands from client applications and executes them
accordingly. The communication between the server and clients is handled
through a \gls{gRPC} interface, which allows for remote procedure calls and
provides a structured way to manage the interactions between the different
components of the system. To note that an undesired, but useful nonetheless,
feature of this architecture is that it allows users to create \textit{bash}
scripts that can interact with the system through the client executable, which
can be useful for quick testing and automation purposes.

% Design choices
It is also relevant to note some of the design choices made during the
implementation of the \gls{SDR} control system. The most impactful one is the
use of \gls{gRPC} for three main reasons: it has proven itself in terms of
reliability and performance, it provides a clear and strongly typed interface
for communication between the server and clients, and allows for reflection,
the latter being one of the most useful feautures during development, as it not
only allows for easy debugging and testing, but also allows the user to easily
discover the available commands and their parameters, which is especially
useful for users that are not familiar with the system. Another quite relevant,
but no as impactful design choice was the use of errors as values instead of
the traditional approach of using exceptions. This aided not only in making a
more concise interface, but also avoids some of the performance pitfalls of
exceptions. \textcolor{Red}{could also talk about xmake instead of cmake, maybe
    the pimpl idiom in device and channel}

% Limitations

\textcolor{Red}{Review}
Some of the limitations of the current implementation of the \gls{SDR} control
system are also related to the client server architecture. One of the main
limitations is that it introduces some latency in the communication between the
server and clients, which can be an issue for real-time applications.
Additionally, whilst on purpose the coupling between the system and \gls{gRPC}
is notable.Another limitation is that the current implementation does not
support all the features of the \gls{SDR} hardware, though this cant really be
considered a problem as those feature are outside the scope of this internship
and can be easily added in the future if needed.

\section{Architecture Diagram}


\usetikzlibrary{positioning, fit, backgrounds, arrows.meta, calc}
 
\begin{tikzpicture}[
  % ── node styles ─────────────────────────────────────────
  comp/.style = {
    draw, rounded corners=3pt,
    minimum width=3.0cm, minimum height=1.5cm,
    text width=2.8cm, align=center,
    font=\small, inner sep=5pt,
  },
  lollipop/.style = {
    draw, circle, minimum size=0.32cm, inner sep=0pt,
    fill=white,
  },
  stereo/.style  = {font=\tiny\itshape, text=gray!60},
  use/.style     = {->, >=open triangle 45, dashed, thick},
  realize/.style = {->, >=open triangle 45, dashed, thick, double distance=1.2pt},
  assoc/.style   = {thick},
  note/.style    = {
    draw=gray!40, fill=yellow!8, rounded corners=2pt,
    font=\tiny, text width=2.8cm, align=left, inner sep=4pt,
  },
  node distance  = 1.4cm,
]
 
% ============================================================
% ROW 1 — CLI Client  |  Server
% ============================================================
\node[comp, fill=green!10, draw=green!55!black]
  (client) at (0,0)
  {\texttt{<<component>>}\\[3pt]\textbf{CLI Client}};
 
\node[comp, fill=green!10, draw=green!55!black,
      right=5.0cm of client]
  (server)
  {\texttt{<<component>>}\\[3pt]\textbf{Server}};
 
% ============================================================
% ROW 2 — Public API  |  dt\_facade  |  dt
% ============================================================
% Public API sits between client and server, one row below
\node[comp, fill=blue!10, draw=blue!50,
      below=3.0cm of $(client)!0.5!(server)$]
  (pubapi)
  {\texttt{<<component>>}\\[3pt]\textbf{Public API}\\[2pt]
   {\tiny\textcolor{gray!60}{(generated by \texttt{protoc})}}};
 
\node[comp, fill=blue!10, draw=blue!50,
      below=3.2cm of server]
  (dtfacade)
  {\texttt{<<component>>}\\[3pt]\textbf{dt\_facade}\\[2pt]
   {\tiny\textcolor{gray!60}{(dtapi high level wrapper)}}};
 
\node[comp, fill=blue!10, draw=blue!50,
      right=2.6cm of dtfacade]
  (dt)
  {\texttt{<<component>>}\\[3pt]\textbf{dt}\\[2pt]
   {\tiny\textcolor{gray!60}{(utility library)}}};
 
% ============================================================
% ROW 3 — DTAPI
% ============================================================
\node[comp, fill=red!10, draw=red!45,
      below=2.8cm of dtfacade]
  (dtapi)
  {\texttt{<<component>>}\\[3pt]\textbf{\gls{DTAPI}}\\[2pt]
   {\tiny\textcolor{gray!60}{(DekTec vendor SDK)}}};
 
% ============================================================
% ROW 4 — Hardware
% ============================================================
\node[comp, fill=red!5, draw=red!30,
      below left=2.2cm and 0.4cm of dtapi]
  (dta2115)
  {\texttt{<<device>>}\\[3pt]\textbf{DTA-2115B}\\[2pt]
   {\tiny\textcolor{gray!60}{PCIe card}}};
 
\node[comp, fill=red!5, draw=red!30,
      below right=2.2cm and 0.4cm of dtapi]
  (dta2116)
  {\texttt{<<device>>}\\[3pt]\textbf{DTA-2116}\\[2pt]
   {\tiny\textcolor{gray!60}{PCIe card}}};
 
% ============================================================
% PROVIDED INTERFACE LOLLIPOPS
% ============================================================
\node[lollipop] (api-lollipop)    at ($(pubapi.north)+(0,0.45)$)  {};
\node[lollipop] (facade-lollipop) at ($(dtfacade.north)+(0,0.45)$){};
\node[lollipop] (dt-lollipop)     at ($(dt.west)+(-0.45,0)$)      {};
\node[lollipop] (dtapi-lollipop)  at ($(dtapi.north)+(0,0.45)$)   {};
 
\draw[assoc] (pubapi.north)   -- (api-lollipop);
\draw[assoc] (dtfacade.north) -- (facade-lollipop);
\draw[assoc] (dt.west)        -- (dt-lollipop);
\draw[assoc] (dtapi.north)    -- (dtapi-lollipop);
 
% ============================================================
% DEPENDENCY ARROWS
% ============================================================
 
% Client  --<<use>>-->  Public API
\draw[use]
  (client.south) .. controls +(0,-1.0) and +(-0.6, 0.6) .. (api-lollipop.west)
  node[stereo, pos=0.45, left=2pt] {<<use>>};
 
% Server  --<<realize>>-->  Public API  (double dashed = realization)
\draw[realize]
  (server.south) .. controls +(0,-0.8) and +(0.6, 0.6) .. (api-lollipop.east)
  node[stereo, pos=0.45, right=2pt] {<<realize>>};
 
% Server  --<<use>>-->  dt_facade
\draw[use]
  (server.south) .. controls +(0,-0.5) and +(0, 0.6) .. (facade-lollipop.north)
  node[stereo, pos=0.5, right=3pt] {<<use>>};
 
% dt_facade  --<<use>>-->  dt
\draw[use]
  (dtfacade.east) -- (dt-lollipop.east)
  node[stereo, midway, above=2pt] {<<use>>};
 
% dt_facade  --<<use>>-->  DTAPI
\draw[use]
  (dtfacade.south) .. controls +(0,-0.6) and +(0.8, 0) .. (dtapi-lollipop.east)
  node[stereo, pos=0.5, right=3pt] {<<use>>};
 
% DTAPI  --[PCIe]-->  hardware
\draw[assoc, gray!55]
  (dtapi.south) .. controls +(0,-0.5) and +(0, 0.5) .. (dta2115.north)
  node[stereo, pos=0.5, left=2pt] {PCIe};
 
\draw[assoc, gray!55]
  (dtapi.south) .. controls +(0,-0.5) and +(0, 0.5) .. (dta2116.north)
  node[stereo, pos=0.5, right=2pt] {PCIe};
 
% ============================================================
% ANNOTATION NOTE
% ============================================================
\node[note, left=0.8cm of pubapi] (note-proto) {
  \textbf{.proto contract}\\
  Client stub \& server skeleton both auto-generated by \texttt{protoc} from the same source.
};
\draw[->, gray!40, dashed, thin] (note-proto.east) -- (pubapi.west);
 
% ============================================================
% BACKGROUND GROUPINGS
% ============================================================
\begin{scope}[on background layer]
  \node[draw=blue!25, dashed, rounded corners=5pt, fill=blue!3,
        fit=(pubapi)(note-proto)(dtfacade)(dt),
        label={[font=\tiny\bfseries, text=blue!45]above right:\gls{gRPC} / Application Layer}]
        (app-group) {};
 
  \node[draw=red!25, dashed, rounded corners=5pt, fill=red!3,
        fit=(dtapi)(dta2115)(dta2116),
        label={[font=\tiny\bfseries, text=red!45]below left:Hardware \& SDK Layer}]
        (hw-group) {};
\end{scope}
 
\end{tikzpicture}