%!TEX root = ../template.tex
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% chapter2.tex
%% NOVA thesis document file
%%
%% Chapter with state of the art
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\typeout{NT FILE chapter2.tex}%

\chapter{State of the Art}
\label{cha:state_of_the_art}

This chapter surveys the existing landscape of \gls{GNSS} signal generation and
simulation tools and related software interfaces, establishing the context in
which the work of this internship was conducted. The goal is to identify the
strengths and limitations of existing solutions and motivate the need for the
developed system.

\section{Commercial GNSS Signal Simulators}

Commercial \gls{GNSS} simulators represent the industry standard for
professional receiver testing and validation. These systems are purpose-built,
highly capable, and widely used in aerospace, automotive, and defence
industries.

The \textbf{Spirent GSS7000} is one of the most prominent examples of a
commercial \gls{GNSS} simulator. It supports simultaneous coherent signal
generation across multiple constellations, including \gls{GPS}, \gls{GLONASS},
\gls{Galileo}, and \gls{BeiDou}, with up to 256 channels flexibly allocated
across constellations \cite{spirentGSS7000}\cite{spirentGSS7000Federal}. The
system is controlled via Spirent's SimGEN software, which provides
comprehensive scenario definition, execution, and data management capabilities.
While powerful, the GSS7000 is a closed, proprietary system with significant
cost, and its control interface is tightly coupled to Spirent's own software
ecosystem, offering limited flexibility for programmatic integration into
custom workflows.

Similar offerings exist from other vendors such as Rohde \& Schwarz
\cite{rohdeSchwarzSMBV100B} and Keysight Technologies \cite{keysightN7609B},
which provide high-fidelity \gls{GNSS} signal generation equipment. These
systems share the same fundamental limitations: they are expensive,
closed-source, and not designed for flexible programmatic control by
third-party software. They are typically operated through vendor-specific
graphical interfaces, making automated testing pipelines difficult to implement
without additional integration effort.

\textcolor{Red}{Can i talk about streamxpress here?}

\section{Open-Source GNSS Signal Generation Tools}

In contrast to commercial solutions, several open-source tools have emerged
that leverage \gls{SDR} hardware to compute and produce \gls{GNSS} baseband
signals at a fraction of the cost. It is important to note that these tools
operate exclusively at the \emph{signal computation} layer: they accept
scenario parameters such as receiver position and satellite ephemeris data, and
produce pre-computed \gls{I/Q} sample files as output. The transmission of
those samples to physical \gls{RF} hardware is outside their scope and must be
handled separately.

\textbf{GPS-SDR-SIM} \cite{gpsSdrSim} is a widely adopted open-source \gls{GPS} signal simulator. It generates \gls{GPS} L1 C/A baseband signal streams as digitised \gls{I/Q} samples, which can then be transmitted via \gls{SDR} hardware such as HackRF or ADALM-Pluto. The tool accepts a user-defined trajectory specified as a CSV file containing \gls{ECEF} positions, or as an NMEA GGA stream, and uses a \gls{GPS} broadcast ephemeris file to compute pseudoranges and Doppler shifts for satellites in view. GPS-SDR-SIM operates in an offline batch mode, producing a pre-computed \gls{I/Q} sample file that must subsequently be fed to a hardware transmitter. It provides no programmatic API for runtime parameter control, and has no built-in support for professional-grade \gls{SDR} hardware.

\textbf{Galileo-SDR-SIM} \cite{galileoSdrSim} extends the GPS-SDR-SIM concept to the Galileo E1B/C signal, generating Galileo baseband signal streams for transmission via \gls{SDR} platforms such as the USRP B210. While it demonstrates the feasibility of open-source multi-constellation signal computation, it shares the same architectural limitation: it is a standalone offline tool that produces \gls{I/Q} sample files, with no structured interface for integration into larger software systems and no support for hardware beyond its target consumer \gls{SDR} platforms.

\textbf{GNSS-SDR} \cite{gnssSdr} is an open-source \gls{GNSS} receiver framework rather than a signal generator. It is designed for signal reception and processing, not generation, and therefore does not address the use case of the present work. Its architecture is nonetheless instructive, demonstrating how modular \gls{SDR}-based \gls{GNSS} software can be structured.

A common limitation across all open-source tools is the absence of any
transmission abstraction layer. These tools solve the signal computation
problem well, but the \gls{I/Q} files they produce have no path to
professional-grade hardware without significant additional engineering effort.
None provide support for DekTec \gls{SDR} cards, nor do they expose a
structured, remotely accessible API for programmatic control of the
transmission process.

\section{Hardware Vendor APIs}

DekTec produces a range of professional \gls{SDR} hardware, including the
DTA-2115 and DTA-2116 PCIe cards used in this internship. These cards are
capable of high-quality, high-bandwidth \gls{RF} signal transmission and are
well suited for professional signal simulation applications.

DekTec provides the \gls{DTAPI}, a C++ library for interfacing with their
hardware. While functional and comprehensive in its coverage of hardware
capabilities, the \gls{DTAPI} is a low-level interface that exposes hardware
primitives directly. Configuring a device requires a detailed understanding of
internal hardware parameters, correct sequencing of API calls, and careful
management of device state. Some tools do exist like TSDuck\cite{ts-duck},
though it does not focus on \gls{GNSS} signal transmission. There are no
existing higher-level abstraction layers for \gls{GNSS} use cases built on top
of the \gls{DTAPI}, at least none that are open-source, nor any interface that
connects the output of \gls{I/Q} generation tools to the hardware transmission
pipeline. Developers must therefore deal with low-level concerns such as FIFO
management, DMA buffer sizing, and modulation flag composition directly, which
increases both development time and the risk of subtle bugs.

\section{Summary and Identified Gap}

Table~\ref{tab:sota_comparison} summarises the key characteristics of the
surveyed solutions.

\begin{table}[htbp]
    \centering
    \caption{Comparison of existing GNSS signal generation and transmission solutions.}
    \label{tab:sota_comparison}
    \begin{tabular}{|l|c|c|c|c|}
        \hline
        \textbf{Solution}  & \textbf{File Playback} & \textbf{Programmatic API} & \textbf{Open Source} & \textbf{DekTec Support} \\
        \hline
        Spirent GSS7000    & N/A                    & Limited                   & \texttimes           & \texttimes              \\
        GPS-SDR-SIM        & \texttimes             & \texttimes                & \checkmark           & \texttimes              \\
        Galileo-SDR-SIM    & \texttimes             & \texttimes                & \checkmark           & \texttimes              \\
        GNSS-SDR           & \texttimes             & \texttimes                & \checkmark           & \texttimes              \\
        DekTec \gls{DTAPI} & \checkmark             & Low-level only            & \texttimes           & \checkmark              \\
        \hline
        \textbf{This work} & \checkmark             & \checkmark                & \texttimes           & \checkmark              \\
        \hline
    \end{tabular}
\end{table}

The survey reveals a clear gap in the existing landscape. Open-source signal
computation tools such as GPS-SDR-SIM produce \gls{I/Q} sample files but
provide no mechanism to transmit them through professional-grade hardware. The
DekTec \gls{DTAPI} provides the necessary hardware access but exposes only
low-level primitives, with no higher-level interface suited to \gls{GNSS}
workflows and no connection to upstream signal generation tools. Commercial
simulators address both layers in a single integrated system, but are closed,
expensive, and not designed for programmatic integration into custom workflows.

The work developed during this internship directly addresses this gap. By
providing a C++ abstraction layer over the \gls{DTAPI}, exposed through a
\gls{gRPC} interface, it bridges the \gls{I/Q} file output of open-source
signal generators and the transmission capabilities of professional DekTec
hardware, enabling programmatic and remotely accessible control of \gls{GNSS}
signal transmission.

\section{Conclusion}
\textcolor{Red}{Remove?}
This chapter has surveyed the state of the art in \gls{GNSS} signal generation
and transmission, covering commercial simulators, open-source signal
computation tools, and hardware vendor APIs. The analysis identified a gap at
the boundary between the \gls{I/Q} generation layer and the hardware
transmission layer: no existing open solution connects these two concerns on
professional-grade hardware through a clean programmatic interface. This gap
motivates the development described in the following chapters.
