% appendices/design_patterns.tex
\chapter{Design Patterns and Architectural Decisions}
\label{cha:design-patterns}

This appendix details the C++ design patterns and architectural decisions
underlying the SDR Control/Interface System and the RS-232 Interface,
expanding on the rationale only briefly mentioned in chapters
\autoref{cha:sdr-control} and \autoref{cha:rs232-interface}.


\section{Errors as Values}
\label{cha:design-patterns:errors}
Exceptions were not used in the SDR Control/Interface System or the RS-232
Interface, as they were deemed to introduce unnecessary overhead and implicit
error contracts. C-style error codes (e.g., returning an integer to indicate success or failure) 
were also avoided, as they may lead to error-prone code due to forgetfulness per part of the developer. 
Instead, the design adopted a more modern approach using \texttt{std::expected<T,E>}, which allows functions to 
return either a successful result of type \texttt{T} or an error of type \texttt{E}, making error handling explicit and predictable, with the benefit of compile-time type checking.
This design choice promotes clearer and more maintainable code, as it forces the caller to handle potential errors explicitly, reducing the likelihood
of unhandled exceptions or overlooked error conditions.
Though it must be noted that this approach can lead to more verbose code at the call site, as it requires explicit checks for success or failure.


\section{Compile-Time Reflection and Serialisation}
\label{cha:design-patterns:reflection}
To facilitate the serialisation and deserialisation of arbitrary C++ structs without requiring manual boilerplate code or annotation macros, the design uses compile-time reflection techniques.
The Boost.PFR library is used for the general case of aggregate types, allowing the system to automatically determine the memory layout of a struct and perform serialisation or deserialisation accordingly.
For non-aggregate types that do not fit the reflection model provided by Boost.PFR (e.g., vectors, maps, or other complex types), a template specialization of a \texttt{CustomSerde<T>} struct is provided as an "escape hatch," allowing developers to define custom serialisation and deserialisation logic for those types.
This approach ensures that the system can handle a wide range of data structures while maintaining flexibility and extensibility, enabling easy adaptation to different command sets as needed.


\section{Protocol Registration as a Type-Erased Command Dispatcher}
\label{cha:design-patterns:registry}
The RS-232 Interface employs a type-erased command dispatcher pattern to manage the registration and handling of different message types. 
This design allows for the dynamic addition of new message types without requiring modifications to the core system, promoting extensibility and maintainability.
Not only does this pattern enable the system to accommodate new message types as they arise, 
but it also decouples the message handling logic from the core processing loop, allowing for cleaner code organization and separation of concerns.



\section{Iterations of the SDR Control System}
\label{cha:design-patterns:sdr-iterations}
 
The \gls{sdr-control-system} described in
\autoref{cha:sdr-control} is the third iteration of the system. The
two preceding attempts are summarised in
\autoref{tab:sdr-iterations}, as the reasoning behind their
abandonment directly informed several of the design decisions
discussed throughout this appendix.
 
\begin{table}[h]
\centering
\caption{Iterations of the SDR Control/Interface System.}
\label{tab:sdr-iterations}
\begin{tabular}{p{2.2cm} p{4.7cm} p{5.8cm}}
\toprule
\textbf{Iteration} & \textbf{Approach} & \textbf{Outcome / Reason for Abandonment} \\
\midrule
1 & 
    Reuse of an existing third-party program for \gls{SDR} control,
    adapted to the \gls{GNSS} use case rather than built from
    scratch. &
    Abandoned due to incompatibilities between the program's
    assumptions 
    (threading model, extensibility of its API) and the requirements of this
    internship, which made adaptation impractical. \\
\addlinespace
2 & A ground-up rewrite aiming for a fully general \gls{SDR} control
    library, supporting an open-ended range of modulation types
    and configurations. &
    Over-generalisation: most of the supported modulation types
    were never required by the actual use case, and the effort
    spent on generality (abstracting over modulation schemes that
    were never exercised) outweighed its benefit, while also
    making the codebase harder to reason about and test. \\
\addlinespace
3 (final) & A ground-up rewrite scoped specifically to the
    DTA-2115B/2116 cards and the \gls{GNSS} \gls{I/Q} transmission
    use case, described in \autoref{cha:sdr-control}. &
    Adopted. Narrowing the scope to the actual requirements
    made the system significantly simpler to implement, test, and
    reason about. \\
\bottomrule
\end{tabular}
\end{table}
 
This outcome is a useful counterpoint to the broader lesson drawn
in \autoref{cha:conclusion-retrospective}, where extensibility and
flexible interfaces are credited with absorbing the unplanned
integration effort described in \autoref{sec:scope-evolution}.
Taken together, the two findings suggest a more nuanced position
than either alone: designing for extensibility paid off where the
\emph{source} of change was external and unpredictable (integration
requirements imposed by GMV's pre-existing systems), but pursuing
generality \emph{pre-emptively}, ahead of any concrete requirement
for it, incurred a cost with no corresponding benefit in the second
iteration. The same tension resurfaces in
\autoref{sec:integration}, where the gRPC layer --- itself a
generalisation in service of a remote-control requirement that the
final integration path did not end up needing --- went unused once
the proprietary system was integrated in-process.
